Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-20 mu/article/article.tex
Starting second Anatoli's review.

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled $Up$, also known as graphone, and the second, one labelled $Alt$, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first one  is by fixing the spin
along a chosen direction, and the resulting spin-velocity is calculated.
Secondly, we choose the spin-velocity direction within {\color{red}(OR along?)}
the surface plane, and calculate {\color{red}(which one?)} the spin-velocity
injection and spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for Up and Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the symmetry Cartesian axis, reaching a
speed of 688\,Km/s with the spin  pointing response is maximized by fixing the
spin-velocity direction along a high at 13$^\circ$ from the surface normal, the
for Up, and 906\,Km/s and the spin pointing at 60$^\circ$ from the surface
normal, for the Alt. These speed values are significantly larger than those of
bulk semiconductors, such as CdSe and GaAs, thus making the hydrogenated
graphene structures excellent candidates for spintronics applications.
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in {\color{red}\st{achieving}} {\color{blue}creating} measurable  the of spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimension (2D)
electronic systems; however, most electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} boron-
nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} noble-
metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in [1]. In this paper, we present two 50\% hydrogenated graphene
noncentrosymmetric structures, both presenting a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13} shown in Fig. \ref{fig:up-struc}. Graphone has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure, we consider here the magnetic isomer of graphone, with the so-called
“chair” structure. In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbonsheet.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/fig1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is also shown.
      structure along with the 
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/fig2}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the $xyz$ directions. The dark (light) spheres are
      the C (H) atoms. The primitive rectangular unit cell is also displayed.}
    \label{fig:alt-struc}
\end{figure}

Using those structures, we address theoretically the spin velocity injection
(SVI) by one-photon absorption of linearly polarized light. Since wehave 2D
structures, we do the analysis for two cases {\color{red}(not clear to me this
sentence)}. The first case is by fixing the spin of the electrons along
vertical to the surface direction {\color{red}(?)}$z$, , with the velocity
directed onto {\color{red}(is this parallel or not?)} the surface of the
structure in the $xy$ plane. In the second case we fix the spin velocity in the
$x$ or $y$ direction and the resulting spin is then directed outward of the
$xy$ plane. The SVI is an optical effect that quantifies the velocity, at which
a PSC moves along the direction $\bf{a}$ with the $\bf{a}$ components to be
determined, with the spin of electron polarized along the direction $\bf{b}$.
polarized light produce an even distribution of electrons in $\mathbf{k}$
space, regardless of the symmetry of the material, resulting in a null
electrical current.\cite{bhatPRL05} Then, the electrons excited to the
conduction bands at opposite $\mathbf{k}$ points will result in opposite spin
polarizations producing no net spin injection in centrosymmetric materials
{\color{red}(?)}.\cite{bhatPRL05} If the crystalline structure of the material
is noncentrosymmetric, the spin polarization injected at a given $\mathbf{k}$
point not necessarily vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88}.
Therefore, since the velocities of electrons at opposite $\mathbf{k}$ points
are opposite, a PSC will be produced. Since both the structures presented here
are noncentrosymmetric, they are good candidates in which this effect can be
induced.

This paper is organized as follows. In Section \ref{sec:theory} we present the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our finding in Section \ref{sec:conclusions}.

@@

where the applied electric field of the laser is given by

@@

%Although, this splitting is small [15,16], is required to obtain the correct
%results\cite{nastosPRB07}. The conduction bands in the noncentrosymmetric
%semiconductors are spin split by a small amount,[15,16] typically smaller than
%the energy width of the laser pulse, and so the pulse excites a coherent
%superposition of the two conduction bands. Even for very long pulses with
%narrow energy widths, dephasing effects lead to an energy width of the bands
%large enough that spin-split states can become quasi-degenerate.
To include the coherence, we follow Ref.~\onlinecite{nastosPRB05} and

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. The spin-
splitting of the valence ($v$) bands is very small, and is neglected throughout

@@

n{\mathbf k}| = 1$, where $n=v,c$, and the fact that $\rho_{vn}({\mathbf
k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. Therefore, using  Eqs. \eqref{z.4}

@@

where repeated Cartesians are summed, and 

@@

\mu^{\mathrm{abcd}}  (\omega) &

@@

\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 

@@

K^{\mathrm{ab}}_{nm}({\mathbf -k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$ which

@@

–identical to Eq. (3) of Bhat et al.\cite{bhatPRL05}, derived using the

@@

We define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $a$, at which the spin moves in a
polarized state along direction $b$. The carrier injection rate $\dot
n(\omega)$ is written as\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k})), 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical response tensor by
$\mathrm{Im} [\epsilon^{\mathrm{a}\mathrm{b}}(\omega)] =
2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

The function ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ allows us to quantify
two very important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$ and calculate the resulting electron velocity. On the other
hand, we can fix the velocity of the electron along $\mathrm{b}$ and study the
resulting direction along which the spin is polarized. To this end, the
additional advantage of  2D structures, besides being noncentrosymmetric, is
that we can use an incoming linearly polarized  light at normal incidence, and
use the  direction of the polarized  electric field to control ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega)$. Indeed, writing ${\mathbf E}(\omega) =
E_0(\omega)(\cos\alpha\,\hat{\mathbf x}+\sin\alpha\,\hat{\mathbf y})$, where
$\alpha$ is the polarization angle, we obtain from Eq. \eqref{eq:vab-w} that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\omega)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{align}
\end{widetext}
for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\omega)=\xi^{\mathrm{yy}}(\omega)\equiv\xi(\omega)$, and
$\xi^{\mathrm{xy}}(\omega)=0$. Next, we identify two options for ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define {\color{red}(calculate?)} the magnitude
of the electron velocity along the plane of the structure, with the spin
polarized along $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular to the
incoming light polarization, respectively. The subscript $\sigma^\mathrm{b}$
denotes the spin along $\mathrm{b}$.

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system's
coordinates defined with a polar angle
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and an azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  vs. $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, starting {\color{red}(going up?)} at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAsThe, shown at the bottom panel . }
\label{fig:vab-str-comp}
\end{figure}

>>>>

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
%\documentclass[preprint,prb,showpacs]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage[pdffitwindow,colorlinks,citecolor={red},linkcolor={blue}]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, one labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first one is by fixing the spin
along a chosen direction, and the resulting SVI is calculated.
Secondly, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt. These speed values are larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

The spin velocity injection (SVI) is an optical effect that quantifies the velocity at which
a PSC moves along the direction $\hat{\mathbf{a}}$, 
with the spin of the electron polarized along the direction $\hat{\mathbf{b}}$.
One photon absorption of polarized
light produces an even distribution of electrons in $\mathbf{k}$
space, regardless of the symmetry of the material, resulting in a null
electrical current.\cite{bhatPRL05} 
Then, the electrons excited to the
conduction bands at opposite $\mathbf{k}$ points will result in opposite spin
polarizations producing no net spin injection in centrosymmetric 
materials.\cite{bhatPRL05} 
If the crystalline structure of the material
is noncentrosymmetric, the spin polarization injected at a given $\mathbf{k}$
point not necessarily vanishes.\cite{alvaradoPRL85,schmiedeskampPRL88}
Therefore, since the velocities of electrons at opposite $\mathbf{k}$ points
are opposite, a PSC will be produced.

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimension (2D)
electronic systems; however, most electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we present two 50\% hydrogenated graphene
noncentrosymmetric structures, both presenting a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

Both the Up and the Alt  structures
are noncentrosymmetric, and therefore,
they are good candidates in which SVI can be
induced. In this article, 
we address theoretically the SVI 
by one-photon absorption of linearly polarized light, analyzing
in our structures,
two possible scenarios of practical interest. 
The first case is by fixing the spin of the electrons along $z$, i.e. 
perpendicular 
to the surface plane, with the resulting velocity
directed along the surface of the
structures in the $xy$ plane. 
In the second case we fix the SVI velocity along the
$x$ or $y$ direction, and then, the resulting spin is directed outward of the
$xy$ plane.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{
Top (top panel) and side (bottom panel) views of the Up
      structure along with the 
      Cartesian $x$, $y$, and $z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
{\color{red}Anatoli: we are confused about the primitive hexagonal
  unit cell shown. For instance, according to us, if we take C$_2$ and
apply the red arrow, it will end up in the site of C$_4$, but these C
atoms according to Table \ref{tab:up-unitcell} are not equivalent, or are they?}} 
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{
Top (top panel) and side (bottom panel) views of the Alt
      structure along with the 
      Cartesian $x$, $y$, and$ z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
}
    \label{fig:alt-struc}
\end{figure}

This paper is organized as follows. In Section \ref{sec:theory} we present the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our findings in Section \ref{sec:conclusions}.

@@

where the applied electric field of the beam of light is given by

@@

To include the coherence, we follow Ref.~\onlinecite{nastosPRB07} and

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. The 
spin-splitting of the valence ($v$) bands is very small, and is neglected throughout

@@

n{\mathbf k}| = 1$, where $n$ goes over all $v$ and $c$ states,
 and the fact that $\rho_{vn}(\mathbf{k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. Therefore, using  Eqs. \eqref{z.4}

@@

where repeated Cartesians upperscripts  are summed, and 

@@

\mu^{\mathrm{abcd}}  (\omega) 

@@

\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 

@@

K^{\mathrm{ab}}_{nm}(-\mathbf{k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$ which

@@

–identical to Eq. (3) of Bhat et al.,\cite{bhatPRL05} derived using the

@@

We define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $\hat{\mathbf{a}}$, at which the spin moves in a
polarized state along direction $\hat{\mathbf{b}}$. 
The carrier injection rate $\dot n(\omega)$ is written as\cite{nastosPRB07}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k})), 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical response tensor by
$\mathrm{Im} [\epsilon^{\mathrm{a}\mathrm{b}}(\omega)] =
2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

The function ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ allows us to quantify
two very important aspects of PSC. On one hand, we can fix the spin direction
along $\hat{\mathbf{b}}$ and calculate the resulting electron velocity. On the other
hand, we can fix the velocity of the electron along $\hat{\mathbf{a}}$
 and study the
resulting direction along which the spin is polarized. To this end, the
additional advantage of  2D structures, besides being noncentrosymmetric, is
that we can use an incoming linearly polarized  light at normal incidence, and
use the  direction of the polarized  electric field to control ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega)$. Indeed, writing ${\mathbf E}(\omega) =
E_0(\omega)(\cos\alpha\,\hat{\mathbf x}+\sin\alpha\,\hat{\mathbf y})$, where
$\alpha$ is the polarization angle, we obtain from Eq. \eqref{eq:vab-w} that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\omega)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{align}
\end{widetext}
since for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\omega)=\xi^{\mathrm{yy}}(\omega)\equiv\xi(\omega)$, and
$\xi^{\mathrm{xy}}(\omega)=0$. Next, we identify two options for ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we calculate the magnitude
of the electron velocity along the plane of the structure, with the spin
polarized along $\hat{\mathbf{b}}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular to the
incoming light polarization direction, respectively. 
The subscript $\sigma^\mathrm{b}$
denotes the spin along $\hat{\mathbf{b}}$.

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system of
coordinates by the defined  polar angle
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and an azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  {\it vs.} $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, becoming different from zero at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}
